Aspects of Dealing with Imperfect Data in Temporal Databases

### 2012-08-27 book/sources/timeDomain.tex
My last (implemented) changes on the chapter.

Data models used by information systems (and in specific, temporal database systems) may represent these underlying time domains using \emph{chronons}~\cite{Dyreson1994}, which can informally be described as the smallest distinguishable time units that can be used in the system. However, to explain what chronons are, an explanation of some other temporal concepts is necessary.

@@

Thus, an instant is basically an instantaneous time point in a time model.

Orthogonal to the classification of underlying time models as linear or cyclic, they can be classified as \emph{discrete}, \emph{dense} or \emph{continuous} models\cite{Dyreson1994}. In a discrete model~\cite{Clifford:1985:AHR:971699.318922}, the notion exists that every instant has a unique successor and the set of instants is seen as a discrete one. Here, intuitively, the set of instants can be seen as isomorphic to the set of natural numbers $\Nat$. In a dense model, the notion exists that between any two instants always lies another. Here, intuitively, the set of instants can be seen as isomorphic to the set of rational numbers $\Q$ (when the set of instants is a discrete one) or the set of real numbers $\R$ (when the set of instants is a continuous one). In a continuous model, the notion also exists that between any two instants always lies another one, but the set of instants is always seen as continuous and there are no ``gaps'' between successive instants.

@@

Data models for time modelling might now represent an underlying time domain using chronons:

@@

A data model may now represent a time model or time domain by a sequence of consecutive chronons. These chronons have identical durations. A data model will typically not specify the exact chronon duration, so it can be fixed later by applications implementing the data model.

The fact that chronons are actually time intervals has a particular effect on the representation of instants and time intervals. In a data model representing a time domain using chronons, an instant is of course represented by a chronon. A time interval may be represented by a set of contiguous chronons, depending on the amount of time the time interval comprises.

@@

\begin{figure}
\centering
\includegraphics[scale=0.5]{graphs/granularityGraph.eps}
\caption{The granularity graph corresponding to example~\ref{examplegrangraph}}
\label{fig:granularity-graph-example}
\end{figure}

@@

Several (groups of) operators have been proposed in order to compare temporal elements and model the temporal relationships between them. Allen~\cite{Allen83} most notably described such relationships between time intervals and as a special case, between instants. Figure~\ref{fig:allen} shows the temporal relationships Allen discerned. Some proposals can be applied to both crisp and other time intervals~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08},~\cite{garrido2009}.

>>>>

Data models used by information systems (and in specific, temporal database systems) may represent an underlying time axis using \emph{chronons}~\cite{Dyreson1994}, which can informally be described as the smallest distinguishable time units that can be used in the system. However, to explain what chronons are, an explanation of some other temporal concepts is necessary.

@@

Thus, an instant is basically an instantaneous time point on the time axis underlying a time model. The term is used in the context of the time model too.

Orthogonal to the classification of time models as linear or cyclic, they can be classified as \emph{discrete}, \emph{dense} or \emph{continuous} models\cite{Dyreson1994},~\cite{Jensen:Dyreson:1998:TheConsensusGlossary}. In a discrete model~\cite{Clifford:1985:AHR:971699.318922}, the notion exists that every instant has a unique successor and the set of (modelled) instants is seen as a discrete one. Here, intuitively, the set of instants can be seen as isomorphic to the set of natural numbers $\Nat$. In a dense model, the notion exists that between any two instants always lies another. Here, intuitively, the set of instants can be seen as isomorphic to the set of rational numbers $\Q$ (when the set of (modelled) instants is a discrete one) or the set of real numbers $\R$ (when the set of (modelled) instants is a continuous one). In a continuous model, the notion also exists that between any two instants always lies another one, but the set of (modelled) instants is always seen as continuous and there are no ``gaps'' between successive instants.

@@

Data models used for time modelling might now represent an underlying time axis using chronons:

@@

A time model contained in a data model may now represent an underlying time axis by a sequence of consecutive chronons. These chronons have identical durations. A data model will typically not specify the exact chronon duration, so it can be fixed later by applications implementing the data model.

The fact that chronons are actually time intervals has a particular effect on the representation of instants and time intervals. In a time model using chronons, an instant is of course represented by a chronon. A time interval may be represented by a set of contiguous chronons, depending on the amount of time the time interval comprises.

@@

\begin{figure}
\centering
\includegraphics[scale=0.5]{graphs/granularityGraph.eps}
\caption{The granularity graph corresponding to example~\ref{examplegrangraph}.}
\label{fig:granularity-graph-example}
\end{figure}

@@

Several (collections of) operators have been proposed in order to compare temporal elements and model the temporal relationships between them. Allen~\cite{Allen83} most notably described such relationships between time intervals and as a special case, between instants. Figure~\ref{fig:allen} shows the temporal relationships Allen discerned. Some proposals can be applied to both crisp and other time intervals~\cite{ohlbach2004},~\cite{nagypal2003},~\cite{schockaert08},~\cite{garrido2009}.
